Default GMSE data structures

### 2020-05-23 notebook/ms/SI3.Rmd
Update the SI files

---
title: 'Example case study in GMSE'
subtitle: 'GMSE: an R package for generalised management strategy evaluation (Supporting Information 3)'
author: A. Bradley Duthie&#xb9;&#xb3;, Jeremy J. Cusack&#xb9;, Isabel L. Jones&#xb9;,
  Jeroen Minderman&#xb9;, Erlend B. Nilsen&#xb2;, Roc&#0237;o A. Pozo&#xb9;, 
  O. Sarobidy Rakotonarivo&#xb9;, Bram Van Moorter&#xb2;, and Nils Bunnefeld&#xb9;
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2] Norwegian Institute for Nature Research, Trondheim, Norway [3] alexander.duthie@stir.ac.uk'
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    reference_docx: docx_template.docx
header-includes:
linestretch: 1
link-citations: yes
linkcolor: blue
bibliography: gmse.bib
biblio-style: apalike
---

@@

```{r, echo = TRUE, eval = FALSE}
sim_1 <- gmse(manager_budget = 10000, user_budget = 10000, res_death_K = 93870, 
              manage_target = 70000, RESOURCE_ini = 35000, plotting = FALSE, 
              stakeholders = 80, land_ownership = TRUE, public_land = 0.4, 
              scaring = FALSE, lambda = 0.275, remove_pr = 0.122, time_max = 30, 
              res_death_type = 3, res_consume = 0.02, res_birth_K = 200000, 
              observe_type = 3, agent_view = 1, converge_crit = 0.01, 
              ga_mingen = 200);
```

```{r, echo = FALSE, eval = TRUE}
load("data/SI3-cull_sim.rda"); # Avoids long compiling time
sim_1 <- sim_cull;
```

@@

```{r, echo = TRUE, fig.height = 6, eval = FALSE}
sim_2 <- gmse(manager_budget = 10000, user_budget = 10000, res_death_K = 93870, 
             manage_target = 70000, RESOURCE_ini = 35000, plotting = FALSE, 
             stakeholders = 80, land_ownership = TRUE, public_land = 0.4, 
             scaring = TRUE, lambda = 0.275, remove_pr = 0.122, time_max = 30, 
             res_death_type = 3, res_consume = 0.02, res_birth_K = 200000, 
             observe_type = 3, agent_view = 1, converge_crit = 0.01, 
             ga_mingen = 200);
```

```{r, echo = FALSE, eval = TRUE}
load("data/SI3-scare_sim.rda"); # Avoids long compiling time
sim_2 <- sim_scare;
```

@@

Within the default sub-models of GMSE, there are several non-default options that can be set using arguments to `gmse` and `gmse_apply`. Brief explanations of these non-default options appear in the [GMSE documentation](https://cran.r-project.org/web/packages/GMSE/GMSE.pdf) (also [listed](https://confoobio.github.io/gmse/reference/gmse.html) on the [GMSE website](https://confoobio.github.io/gmse/)). Here we further explain some of the less obvious or less intuitive options available in GMSE v0.4.0.7, for which additional explanation beyond the GMSE documentation may be warranted for application to case studies such as the Taiga Bean Geese case study example demonstrated above. Future versions of GMSE are expected to expand upon these options.

**res_move_type**: Resources in default GMSE v0.4.0.7 sub-models can move on the landscape in one of three ways. These different movement types are listed below, and are taken as arguments to `gmse` (e.g., `gmse(res_move_type = 2)`).

@@

**res_death_type**: Resources in GMSE v0.4.0.7 can be removed from the model in one of three ways within the default resource sub-model, thereby simulating resource death. It is important to note that `res_death_type` refers only to removal that occurs within the resource sub-model; removal from user actions (e.g., death from shooting) occurs within the user sub-model.

  1. *Density-independent removal*. Given this option, there is no density-dependent removal of resources within the resource sub-model. Resources are simply removed with a probability of `removal_pr` each time the sub-model is run. This option should be used with great care because in the absence of density-dependence, it is possible for the population of resources to increase without limit and thereby greatly slow down simulation times. 
  2. *Density-dependent removal* (default). Given this option, removal of resources within the resource sub-model is caused entirely by density effects. The probability that a resource is removed is a function of the resource carrying capacity set using the `res_death_K` parameter. If the number of resources ($N$) is greater than `res_death_K` ($K$), then the probability that an individual resource is removed is defined by $(N - K) / N$. Removal then occurs independently for each resource based on this probability.
  3. *Density-independent and density-dependent removal*. This option allows a combination of both options (1) and (2), which affect removal independently. In other words, each resource is assigned a probability of removal `removal_pr` as in (1), and then independently assigned another probability of removal based on `res_death_K` as in (2). If removal occurs as a consequence of (1), (2), or both, then the resource is removed.

@@

**obs_move_type**: Observation of agents (manager and users) in default GMSE v0.4.0.7 sub-models can occur in the same four ways as allowed in `res_move_type` (default `obs_move_type = 1`), with movement distance defined by the parameter `agent_move` (default 50). Currently, the only effect of this movement is to allow for spatially autocorrelated sampling of resources by the manager in the observation model. If the manager is using density-based sampling (`observe_type = 0`) and sampling more than one subset of the landscape (`times_observe > 1`), then the manager's movement between subsets is governed by `res_move_type` (movement rules) and `agent_move` (movement distance). Hence, to have the manager observe the population many times in one area of the landscape, the `times_observe` option should be increased above one and `agent_move` should be decreased to a low value.

>>>>

---
title: 'Example case study in GMSE'
subtitle: 'GMSE: an R package for generalised management strategy evaluation (Supporting Information 3)'
author: |
        | A. Bradley Duthie&#xb9;&#xb3;, Jeremy J. Cusack&#xb9;, Isabel L. Jones&#xb9;, Jeroen Minderman&#xb9;, 
        | Erlend B. Nilsen&#xb2;, Roc&#0237;o A. Pozo&#xb9;, O. Sarobidy Rakotonarivo&#xb9;, 
        | Bram Van Moorter&#xb2;, and Nils Bunnefeld&#xb9;
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2] Norwegian Institute for Nature Research, Trondheim, Norway [3] alexander.duthie@stir.ac.uk'
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    reference_docx: docx_template.docx
header-includes:
linestretch: 1
link-citations: yes
linkcolor: blue
bibliography: gmse.bib
biblio-style: apalike
---

@@

```{r, echo = TRUE, eval = TRUE, results = 'hide'}
sim_1 <- gmse(manager_budget = 10000, user_budget = 10000, res_death_K = 93870, 
              manage_target = 70000, RESOURCE_ini = 35000, plotting = FALSE, 
              stakeholders = 80, land_ownership = TRUE, public_land = 0.4, 
              scaring = FALSE, lambda = 0.275, remove_pr = 0.122, time_max = 30, 
              res_death_type = 3, res_consume = 0.02, res_birth_K = 200000, 
              observe_type = 3, agent_view = 1, converge_crit = 0.01, 
              ga_mingen = 200);
```

```{r, echo = FALSE, eval = TRUE}
# load("data/SI3-cull_sim.rda"); # Avoids long compiling time
# sim_1 <- sim_cull;
```

@@

```{r, echo = TRUE, fig.height = 6, eval = TRUE, results = 'hide'}
sim_2 <- gmse(manager_budget = 10000, user_budget = 10000, res_death_K = 93870, 
             manage_target = 70000, RESOURCE_ini = 35000, plotting = FALSE, 
             stakeholders = 80, land_ownership = TRUE, public_land = 0.4, 
             scaring = TRUE, lambda = 0.275, remove_pr = 0.122, time_max = 30, 
             res_death_type = 3, res_consume = 0.02, res_birth_K = 200000, 
             observe_type = 3, agent_view = 1, converge_crit = 0.01, 
             ga_mingen = 200);
```

```{r, echo = FALSE, eval = TRUE}
# load("data/SI3-scare_sim.rda"); # Avoids long compiling time
# sim_2 <- sim_scare;
```

@@

Within the default sub-models of GMSE, there are several non-default options that can be set using arguments to `gmse` and `gmse_apply`. Brief explanations of these non-default options appear in the [GMSE documentation](https://cran.r-project.org/web/packages/GMSE/GMSE.pdf) (also [listed](https://confoobio.github.io/gmse/reference/gmse.html) on the [GMSE website](https://confoobio.github.io/gmse/)). Here we further explain some of the less obvious or less intuitive options available in GMSE, for which additional explanation beyond the GMSE documentation may be warranted for application to case studies such as the Taiga Bean Geese case study example demonstrated above. Future versions of GMSE are expected to expand upon these options.

**res_move_type**: Resources in default GMSE sub-models can move on the landscape in one of three ways. These different movement types are listed below, and are taken as arguments to `gmse` (e.g., `gmse(res_move_type = 2)`).

@@

**res_birth_type**: Resources in GMSE can be added to the model in one of two ways within the default resource sub-model, thereby simulating resource birth. It is important to note that `res_birth_type` refers only to additions that occur within the resource sub-model; additions from user actions (e.g., the `help_offspring` option) occurs within the user sub-model.

  0. *No explicit resource addition*. This option does not cause any baseline resource additions, but resource birth is still possible if `consume_repr > 0`. Under such conditions, landscape yield consumed by the resource will result in `floor(yield_consumed / consume_repr)` offspring produced. The amount of yield a resource consumes will be affected by both `times_feeding` (how many times the resource moves and feeds on a new landscape cell during a time step) and `res_consume` (the proportion of yield on a cell that is is consumed upon visit to the cell). Note that `consume_repr` can still be used when `res_birth_type > 0` (below), resulting in multiple sources of resource addition.
  1. *Fixed number added per resource*. For this option, each resource produces `lambda` new resources in each time step.
  2. *Poisson sampling*. For this option, each resource produces a number of offspring sampled independently from a Poisson distribution in which `lambda` is the rate parameter (i.e., each resource is expected to produce `lambda` offspring).

**res_death_type**: Resources in GMSE can be removed from the model in one of four ways within the default resource sub-model, thereby simulating resource death. It is important to note that `res_death_type` refers only to removal that occurs within the resource sub-model; removal from user actions (e.g., death from shooting) occurs within the user sub-model.

  0. *No explicit density independent or dependent removal*. This option does not cause removal to be *directly* caused in a density independent or dependent way. Nevertheless, removal may still be possible if the argument `consume_surv > 0` (its default value is 0). In this case, survival of a resource in a time step depends on that resource consuming at least `consume_surv` in yield on the landscape. In such cases, a natural carrying capacity will be generated as more resources result in higher total consumption. The amount yield that a resource consumes is further affected by both `times_feeding` (how many times the resource moves and feeds on a new landscape cell during a time step) and `res_consume` (the proportion of yield on a cell that is is consumed upon visit to the cell). Note that `consume_surv` is can still be used when `res_death_type > 0` (below), resulting in multiple potential sources of removal.
  1. *Density-independent removal*. Given this option, there is no density-dependent removal of resources within the resource sub-model. Resources are simply removed with a probability of `removal_pr` each time the sub-model is run. This option should be used with great care because in the absence of density-dependence, it is possible for the population of resources to increase without limit and thereby greatly slow down simulation times. 
  2. *Density-dependent removal* (default). Given this option, removal of resources within the resource sub-model is caused entirely by density effects. The probability that a resource is removed is a function of the resource carrying capacity set using the `res_death_K` parameter. If the number of resources ($N$) is greater than `res_death_K` ($K$), then the probability that an individual resource is removed is defined by $(N - K) / N$. Removal then occurs independently for each resource based on this probability.
  3. *Density-independent and density-dependent removal*. This option allows a combination of both options (1) and (2), which affect removal independently. In other words, each resource is assigned a probability of removal `removal_pr` as in (1), and then independently assigned another probability of removal based on `res_death_K` as in (2). If removal occurs as a consequence of (1), (2), or both, then the resource is removed.

@@

**obs_move_type**: Observation of agents (manager and users) in default GMSE sub-models can occur in the same four ways as allowed in `res_move_type` (default `obs_move_type = 1`), with movement distance defined by the parameter `agent_move` (default 50). Currently, the only effect of this movement is to allow for spatially autocorrelated sampling of resources by the manager in the observation model. If the manager is using density-based sampling (`observe_type = 0`) and sampling more than one subset of the landscape (`times_observe > 1`), then the manager's movement between subsets is governed by `res_move_type` (movement rules) and `agent_move` (movement distance). Hence, to have the manager observe the population many times in one area of the landscape, the `times_observe` option should be increased above one and `agent_move` should be decreased to a low value.

### 2020-05-23 notebook/ms/SI4.Rmd
Update the SI files

---
title: 'Advanced case study options'
subtitle: 'GMSE: an R package for generalised management strategy evaluation (Supporting Information 4)'
author: A. Bradley Duthie&#xb9;&#xb3;, Jeremy J. Cusack&#xb9;, Isabel L. Jones&#xb9;,
  Jeroen Minderman&#xb9;, Erlend B. Nilsen&#xb2;, Roc&#0237;o A. Pozo&#xb9;, 
  O. Sarobidy Rakotonarivo&#xb9;, Bram Van Moorter&#xb2;, and Nils Bunnefeld&#xb9;
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2] Norwegian Institute for Nature Research, Trondheim, Norway [3] alexander.duthie@stir.ac.uk'
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    reference_docx: docx_template.docx
header-includes: 
linestretch: 1
link-citations: yes
linkcolor: blue
bibliography: gmse.bib
biblio-style: apalike
---

>>>>

---
title: 'Advanced case study options'
subtitle: 'GMSE: an R package for generalised management strategy evaluation (Supporting Information 4)'
author: A. Bradley Duthie&#xb9;&#xb3;, Jeremy J. Cusack&#xb9;, Isabel L. Jones&#xb9;,
  Jeroen Minderman&#xb9;, Erlend B. Nilsen&#xb2;, Roc&#0237;o A. Pozo&#xb9;, 
  O. Sarobidy Rakotonarivo&#xb9;, Bram Van Moorter&#xb2;, and Nils Bunnefeld&#xb9;
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2] Norwegian Institute for Nature Research, Trondheim, Norway [3] alexander.duthie@stir.ac.uk'
output:
  html_document: default
  pdf_document:
    citation_package: natbib
    fig_caption: yes
  word_document:
    fig_caption: yes
    reference_docx: docx_template.docx
header-includes: 
linestretch: 1
link-citations: yes
linkcolor: blue
bibliography: gmse.bib
biblio-style: apalike
---

### 2020-05-23 notebook/ms/SI7.Rmd
Some proposed subheading additions in SI7
Update the SI files

---
title: "Default GMSE data structures"
author: A. Bradley Duthie&#xb9;&#xb3;, Jeremy J. Cusack&#xb9;, Isabel L. Jones&#xb9;,
  Jeroen Minderman&#xb9;, Erlend B. Nilsen&#xb2;, Roc&#0237;o A. Pozo&#xb9;, O. Sarobidy
  Rakotonarivo&#xb9;, Bram Van Moorter&#xb2;, and Nils Bunnefeld&#xb9;
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2] Norwegian Institute for Nature Research, Trondheim, Norway [3] alexander.duthie@stir.ac.uk'
output:
  html_document: default
  pdf_document:
    citation_package: natbib
    fig_caption: yes
  word_document:
    fig_caption: yes
    reference_docx: docx_template.docx
header-includes: 
link-citations: yes
linkcolor: blue
bibliography: gmse.bib
subtitle: 'GMSE: an R package for generalised management strategy evaluation (Supporting
  Information 7)'
biblio-style: apalike
---

>>>>

---
title: "Default GMSE data structures"
author: |
        | A. Bradley Duthie&#xb9;&#xb3;, Jeremy J. Cusack&#xb9;, Isabel L. Jones&#xb9;, Jeroen Minderman&#xb9;, 
        | Erlend B. Nilsen&#xb2;, Roc&#0237;o A. Pozo&#xb9;, O. Sarobidy Rakotonarivo&#xb9;, 
        | Bram Van Moorter&#xb2;, and Nils Bunnefeld&#xb9;'
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2] Norwegian Institute for Nature Research, Trondheim, Norway [3] alexander.duthie@stir.ac.uk'
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    reference_docx: docx_template.docx
header-includes: 
link-citations: yes
linkcolor: blue
bibliography: gmse.bib
subtitle: 'GMSE: an R package for generalised management strategy evaluation (Supporting
  Information 7)'
biblio-style: apalike
---
